### 2023-05-15 paper/markdown/methods_figures_tables.Rmd

---
title: "Persistent low-grade inflammation, SARS-CoV-2 infection and mental health deterioration signs impact on systemic levels of aminoacid neurotransmitter precursors"
subtitle: "Manuscript parts"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::word_document2:   
    reference_docx: ms_template.docx

bibliography: infl_biblio.bib

csl: frontiers-in-immunology.csl

header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)
library(flextable)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      dev = "png", dpi = 600, 
                      tab.cap.pre = 'Supplementary Table ', tab.cap.sep = " ", 
                      tab.cap.style = 'Table Heading')

set_flextable_defaults(font.family = 'Cambria', font.size = 10)

```

\newpage

# Methods

## Bioinformatic and statistical analysis

R version 4.2.0 was employed for the data analysis.

Normalized, age- and sex-adjusted, log~2~-transformed serum protein and metabolite levels and clinical data for the INCOV cohort were extracted from the report by Su at al. [@Su2022].

Normality and homogeneity of variances was investigated by Shapiro-Wilk and Levene test, respectively. 
Non-parametric tests or normality/homogeneity-improving variable transformations (logarithm and square root) were used in cases of non-normal distribution. 
Comparison of serum metabolite and cytokine levels between uninfected controls and COVID-19 individuals at consecutive timepoints after infection was done by Kruskal-Wallis test with Benjamini-Hochberg-corrected Mann-Whitney post-hoc test [@Benjamini1995]. 
Correlations of metabolite and cytokine serum levels and of mental disorder scoring with metabolite and cytokine levels was investigated by Spearman test. 
Correlation of SIMMUN dataset metabolite levels with age was accomplished by Pearson test. 
Significance of differences in SIMMUN cohort metabolite and inflammatory marker serum concentrations between participants stratified by gender or SARS-CoV-2 was assessed by two-tailed T test.

In multi-parameter linear modeling of aminoacid neurotransmitter precursors and their decay products, neopterin (representative inflammation marker), SARS-CoV-2 infection status, scores of anxiety (HADS), depression (HADS) and stress (PSS-4), age and gender served as candidate explanatory variables. 
Modeling responses and numeric explanatory variables were normalized. 
The models models with the complete explanatory variable set were optimized by Bayesian information criterion (BIC) driven backwards elimination of non-significant terms. 
The normality and homogeneity model residual assumptions were by Shapiro-Wilk and Levene test, respectively, and additionally visually inspected in standard diagnostic plots (residuals versus fitted, quantile-quantile plots). 
Reproducibility of the optimized multi-parameter models was investigated by repeated cross-validation (50 repeats, 10 folds) and by comparison of the RMSE and $R^2$ statistics obtained with the training dataset and in cross-validation.

Details of bioinformatic and statistical analysis are provided in __Supplementary Methods__.

\newpage

# Results

\newpage

# Discussion

\newpage

## Data and code availability

Anonymized patient data will be made available upon request to the corresponding author. 
The INCOV cohort data are publicly available.
The study pipeline code is available at https://github.com/PiotrTymoszuk/stigma_validation.

\newpage

# Tables

```{r tab-stigma-cohort, tab.cap = 'Characteristic of the local SIMMUN cohort stratified by SARS-CoV-2 infection status. Numeric variables are presented as medians with interquartile ranges. Categorical variables are presented as percentages and counts within complete observations.'}

flextable::flextable(tabs$stigma_cohort) %>% 
  width(1, width = 4, unit = 'cm') %>% 
  width(2:3, width = 4, unit = 'cm') %>% 
  width(4:5, width = 2.8, unit = 'cm') %>% 
  footnote(4, 1, 
           value = as_paragraph('normal: body mass index [BMI] < 25 kg/mm\u00B2, overweight: BMI 25 - 30 kg/mm\u00B2, obesity: BMI > 30 kg/mm\u00B2'), 
           part = 'body', 
           ref_symbols = 'a') %>% 
  footnote(7:9, 1, 
           value = as_paragraph('HADS: hospital anxiety and depression scale'), 
           part = 'body', 
           ref_symbols = 'b') %>% 
  footnote(10, 1, 
           value = as_paragraph('PSS-4: 4 item perceived stress scale'),
           part = 'body', 
           ref_symbols = 'c') %>%
  footnote(11, 1, 
           value = as_paragraph('immunoglobulin gamma anti receptor binding domain of the SARS-CoV-2 S1 protein, arbitrary units (AU)'), 
           part = 'body', 
           ref_symbols = 'd') %>% 
  footnote(12, 1, 
           value = as_paragraph('mild: ambulatory care, moderate: hospitalized, normal ward, severe: intensive care unit'), 
           part = 'body', 
           ref_symbols = 'e') %>% 
  theme_vanilla

```

\newpage

```{r tab-incov-cohort, tab.cap = 'Characteristic of the external INCOV cohort. Numeric variables are presented as medians with interquartile ranges. Categorical variables are presented as percentages and counts within complete observations.'}

flextable::flextable(tabs$incov_cohort) %>% 
  width(1, width = 4, unit = 'cm') %>% 
  width(2:3, width = 4, unit = 'cm') %>% 
  width(4:5, width = 2.8, unit = 'cm') %>% 
  footnote(3, 1, 
           value = as_paragraph('normal: body mass index [BMI] < 25 kg/mm\u00B2, overweight: BMI 25 - 30 kg/mm\u00B2, obesity: BMI > 30 kg/mm\u00B2'), 
           part = 'body', 
           ref_symbols = 'a') %>% 
  footnote(5, 1, 
           value = as_paragraph('WHO ordinal scale for clinical improvement; mild: 1 - 2, moderate: 3 - 4, severe: 5 - 6, critical: 7'), 
           part = 'body', 
           ref_symbols = 'b') %>% 
  theme_vanilla

```

\newpage

# Figures

```{r fig-incov-neuro-tc, fig.width = mm_inch(figs$incov_neuro_tc$w), fig.height = mm_inch(figs$incov_neuro_tc$h), fig.cap = 'Serum levels of neurotransmitter precursor and their decay products in course of COVID-19 and recovery, INCOV cohort.'}

figs$incov_neuro_tc$plot

```

__Figure \@ref(fig:fig-incov-neuro-tc). Serum levels of neurotransmitter precursor and their decay products in course of COVID-19 and recovery, INCOV cohort.__

_Serum levels of tryptophan (TRP), kynurenine (KYN), quinolinic acid (QUIN), phenylalanine (PHE) and tyrosine (TYR) in serum of uninfected controls and COVID-19 individuals during acute, sub-acute and recovery phase of the disease in the INCOV cohort. Statistical significance was determined by Kruskal-Wallis test with Benjamini-Hochberg-corrected Mann-Whitney U test. Normalized serum level concentrations are presented in box plots. Points represent single samples. The Kruskal-Wallis test results are indicated in the plot captions. Results of the post-hoc tests are indicated in the plots. Numbers of complete observations are displayed under the plots._

\newpage

```{r fig-met-infl, fig.width = mm_inch(figs$met_infl$w), fig.height = mm_inch(figs$met_infl$h), fig.cap = 'Correlation of serum levels of aminoacid precursors of neurotransmitters and their decay products with markers of inflammation.'}

figs$met_infl$plot

```

__Figure \@ref(fig:fig-met-infl). Correlation of serum levels of aminoacid precursors of neurotransmitters and their decay products with markers of inflammation.__

_Serum levels of tryptophan (TRP), kynurenine (KYN), quinolinic acid (QUIN), kynurenine/tryptophan ratio (KYN/TRP), phenylalanine (PHE), tyrosine (TYR) and phenylalanine/tyrosine ratio (PHE/TYR) were correlated with serum levels of inflammatory markers interleukin 6 (IL6), interleukin 10 (IL10), tumor necrosis factor-alpha (TNF), interferon-gamma (INFG), C-reactive protein (CRP), neopterin (NEO) and neutrophil/lymphocyte ratio (NLR) in uninfected and SARS-CoV-2-infected individuals from the SIMMUN (A) and INCOV cohort (B). Statistical significance was assessed by Spearman test. Correlation coefficients are presented in bubble plots. Point sizes correspond to absolute values of correlation coefficient. Point color corresponds to the correlation coefficient value. Correlation coefficients for significant effects are highlighted in bold._

\newpage

```{r fig-cov, fig.width = mm_inch(figs$cov$w), fig.height = mm_inch(figs$cov$h), fig.cap = 'Association of neurotransmitter precursor aminoacids and their decay products with SARS-CoV-2 infection and anxiety, depression and stress scoring in the SIMMUN cohort.'}

figs$cov$plot

```

__Figure \@ref(fig:fig-cov). Association of neurotransmitter precursor aminoacids and their decay products with SARS-CoV-2 infection and anxiety, depression and stress scoring in the SIMMUN cohort.__

_(A) Serum levels of tryptophan (TRP), kynurenine (KYN), kynurenine/tryptophan ratio (KYN/TRP), phenylalanine (PHE), tyrosine (TYR) and phenylalanine/tyrosine ratio (PHE/TYR) in the SIMMUN study participants stratified by SARS-CoV-2 infections status. Statistical significance was determined by two-tailed T test. Serum levels are shown in box plots. Points represent single observations. Significance p values are displayed in the plot captions, numbers of complete observations are indicated in the plot axes._

_(B) Serum levels of TRP, KYN, KYN/TRP, PHE, TYR and PHE/TYR were correlated with anxiety, depression and stress scores in the SIMMUN study participants (n = `r nrow(stigma$data)`). Statistical significance was assessed by Spearman test. Correlation coefficients are presented in bubble plots. Point sizes correspond to absolute values of correlation coefficient. Point color corresponds to the correlation coefficient value. Correlation coefficients for significant effects are highlighted in bold._

\newpage

```{r fig-multi-model, fig.width = mm_inch(figs$multi_model$w), fig.height = mm_inch(figs$multi_model$h), fig.cap = 'Results of multi-parameter modeling of aminoacid neurotransmitter precursor and their decay products.'}

figs$multi_model$plot

```

__Figure \@ref(fig:fig-multi-model). Results of multi-parameter modeling of aminoacid neurotransmitter precursor and their decay products.__

_Effects of systemic inflammation (neopterin, NEO), SARS-CoV-2 infection status, titre of immunoglobulin gamma against the receptor binding domain of the S1 SARS-CoV-2 protein (anti-RBD IgG), scores of anxiety, depression and stress, age and sex was investigated by multi-parameter linear regression with backward elimination of non-significant terms. Overall model validity was assessed by likelihood-ratio test (LRT). Significant model coefficient estimates with 95% confidence intervals are presented in Forest plots. LRT results (F values, degrees of freedom and p values) and total numbers of complete observations are indicated in the plot captions. Model intercepts (baseline) are indicated in the X axis titles._

\newpage

```{r fig-summary, fig.width = mm_inch(figs$summary$w), fig.height = mm_inch(figs$summary$h), fig.cap = 'Schematic representation of effects of inflammation, SARS-CoV-2 and depression/anxiety symptoms on indolamine and catecholamine neurotransmitter precursor metabolism.'}

figs$summary$plot

```

__Figure \@ref(fig:fig-summary). Schematic representation of effects of inflammation, SARS-CoV-2 and depression/anxiety symptoms on indolamine neurotransmitter precursor metabolism.__

_Effects on metabolism of tryptophan (TRP), the serotonin precursor. SARS-CoV-2-dependent and -independent inflammation reduces systemic TRP levels, stimulates IDO1-mediated conversion of TRP to kynurenin (KYN) and quinolinic acid (QUIN). Independently of inflammation, during recovery from SARS-CoV-2 infection, TRP/KYN/QUIN conversion is stimulated as well. As a result, less substrate for production of the anti-depressive and anxiolytic neurotransmitter serotonin is available, which may predispose to depression or anxiety. Depression and anxiety symptoms on their own are paralleled by decrease in circulating TRP._

\newpage

# Acknowledgements

The scheme of report findings was created with bioicons by Simon Duerr. 
The SARS-CoV-2 virus image was created by Hanna Vega and licensed under the terms of a Creative Commons CC-BY SA 4.0

# References