### 2023-05-15 paper/markdown/supplementary_material_legacy.Rmd

---
title: "Persistent low-grade inflammation, SARS-CoV-2 infection and mental health deterioration signs impact on systemic levels of aminoacid neurotransmitter precursors"
subtitle: "Supplementary Material"
author: "PSY Team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::word_document2:   
    reference_docx: ms_template.docx

bibliography: infl_biblio.bib
csl: frontiers-in-immunology.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      dev = "png", dpi = 600, 
                      tab.cap.pre = 'Supplementary Table ', tab.cap.sep = " ", 
                      tab.cap.style = 'Table Heading')

set_flextable_defaults(font.family = 'Cambria', font.size = 10)

```

\newpage

# Supplementary Methods

## Software

Proteome and metabolome data were analyzed with R version 4.2.0. 
General data transformation tasks were accomplished with the _tidyverse_ package bundle [@Wickham2019], _rlang_ [@Henry2022] and the development package _trafo_ (https://github.com/PiotrTymoszuk/trafo). 
Statistical data testing was done with the packages _rstatix_ [@Kassambara2021], _ggpubr_ [@Kassambara2020] along with the development package _ExDA_ (https://github.com/PiotrTymoszuk/ExDA). 
In linear modeling, base R functions, the development packages _lmqc_ (https://github.com/PiotrTymoszuk/lmqc) and _caretExtra_ (https://github.com/PiotrTymoszuk/caretExtra), and the package _caret_ [@Kuhn2008] were utilized.

Results were visualized with tools provided by the packages _ggplot2_ [@Wickham2016], _cowplot_ [@Wilke2019] and _ExDA_. 
Manuscript and supplementary tables were created with _flextable_ [@Gohel2022]. 
Supplementary Material file was written in the _rmarkdown_ environment [@Allaire2022] and rendered with the _knitr_ [@Xie2022] and _bookdown_ [@Xie2016] packages.

## INCOV cohort data import and transformation

Proteome and metabolome data in form of normalized, age- and sex-adjusted, log~2~-transformed serum levels as well as clinical information (sex, SARS-CoV-2 infection status, COVID-19 severity, timepoint, post-COVID-19 syndrome status and particular symptoms) for the INCOV cohort were extracted from supplementary tables accompanying the report by Su and colleagues [@Su2022].

Characteristic of the INCOV cohorts is presented in __Table 2__. 
Numbers of available INCOV cohort samples and the sampling timepoints are shown in __Supplementary Table S\@ref(tab:tab-samples)__.

## Variable distribution and transformation

Distribution normality and variance homogeneity of normalized cytokine and metabolite serum levels was assessed by Shapiro-Wilk and Levene test, respectively. 
The distribution testing revealed substantial deviations from normality for multiple study parameters. 
For this reason statistical hypothesis testing in the INCOV dataset was done with non-parametric tests. 
For SIMMUN cohort metabolite and inflammation marker levels (KYN, KYN/TRP ratio, PHE, TYR and PHE/TYR ratio, NEO, CRP, IL6, anti-RBD IgG) used in analysis of correlation with age and mental disorder scoring, comparison between participants stratified by SARS-CoV-2 infection status or gender as well as for linear modeling, logarithm and square root transformations was used, which improved both normality and variance homogeneity.

## Statistical hypothesis testing

Comparison of normalized serum cytokine and metabolite levels between uninfected controls and COVID-19 individuals from the INCOV cohort at consecutive timepoints after symptom onset was done with Kruskal-Wallis test. 
Differences between SARS-CoV-2-negative controls and consecutive timepoints of COVID-19 were investigated with Mann-Whitney post-hoc U test corrected for multiple testing with Benjamini-Hochberg method [@Benjamini1995].

Correlation of serum cytokine and metabolite levels in uninfected controls and COVID-19 individuals at consecutive timepoints in the INCOV and SIMMUN datasets was analyzed by Spearman test. 
Correlation of serum cytokine and metabolite levels with scores of anxiety, depression and stress was assessed by Spearman test. 
Correlation of SIMMUN dataset metabolite levels with age was assessed by Pearson test.

Comparison of SIMMUN dataset metabolite and inflammatory marker levels between participants stratified by SARS-CoV-2 infection status or gender was done with two-tailed T test.

## Multi-parameter linear regression with backwards elimination

Effects of the representative inflammation marker (NEO), SARS-CoV-2 infection status, titre of immunoglobulin gamma against the receptor binding domain of the S1 SARS-CoV-2 protein (anti-RBD IgG), anxiety (HADS), depression (HADS) and stress scoring (PSS-4), age and gender on systemic levels of aminoacid neurotransmitter precursor and their decay products (TRP, KYN, KYN/TRP ratio, PHE, TYR and PHE/TYR ratio) was assessed by multi-parameter linear regression with backwards elimination. 
Modeling responses and numeric explanatory variables were transformed with log or squared root functions to improve normality and normalized.

Full models including the complete set of explanatory variables listed above were constructed (function `make_lm()`, package _lmqc_) and optimized by Bayesian information criterion (AIC) driven backwards elimination of non-significant terms (method `step()`, package _lmqc_). 
Normality and homogeneity of distribution of the model residuals was checked by Shapiro-Wilk and Levene test, respectively (method `summary(type = 'assumptions')`, package _lmqc_) and visually inspected in standard diagnostic plots of model residuals (residuals vs fitted, quantile-quantile plots, method `plot()` called for the model objects). 
Fit stats ($R^2$ and root mean squared error [RMSE]) were retrieved from the model objects by calling `summary(type = 'fit')` (package _lmqc_). 
Validity of the optimized models was determined by likelihood-ratio test (LRT) versus the respective null models (method `anova()`). 
Reproducibility and proper parameterization of the optimized multi-parameter models was investigated by repeated cross-validation (50 repeats, 10 folds, function `train(method = 'lm')`, package _caret_) and by comparison of the RMSE and $R^2$ statistics obtained with in the training dataset and in cross-validation (method `summary()`, package _caretExtra_). 
As presented in __Supplementary Figure S\@ref(fig:fig-fit-stats)A__, similar values of error fit in the training and cross-validation data suggest good reproducibility of the optimized models and lack of over-parameterization.

## Data and code availability

Anonymized local cohort data will be made available upon request to the corresponding author. 
The INCOV cohort data are publicly available.
The study pipeline code is available at https://github.com/PiotrTymoszuk/SIMMUN_validation.

\newpage

# Supplementary Tables

```{r tab-samples, tab.cap = 'Number of available samples and sampling timepoints in the INCOV cohort.'}

flextable::flextable(suppl_tabs$samples) %>% 
  width(1, width = 4, unit = 'cm') %>% 
  width(2, width = 6, unit = 'cm') %>% 
  width(3, width = 4, unit = 'cm') %>% 
  footnote(1, 2, value = as_paragraph('Median with interquartile range, days after symptom onset.'), ref_symbols = 'a', part = 'header') %>% 
  theme_vanilla

```

\newpage

# Supplementary Figures

```{r fig-incov-neuro-tc, fig.width = mm_inch(suppl$incov_infl_tc$w), fig.height = mm_inch(suppl$incov_infl_tc$h), fig.cap = 'Serum levels of inflammatory cytokines in course of COVID-19 and recovery, INCOV cohort.'}

suppl$incov_infl_tc$plot

```

__Supplementary Figure S\@ref(fig:fig-incov-neuro-tc). Serum levels of inflammatory cytokines in course of COVID-19 and recovery, INCOV cohort.__

_Serum levels of interleukin 6 (IL6), interleukin 10 (IL10), tumor necrosis factor alpha (TNF) and interferon-gamma (IFN) in serum of uninfected controls and COVID-19 individuals during acute, sub-acute and recovery phase of the disease in the INCOV cohort. Statistical significance was determined by Kruskal-Wallis test with Benjamini-Hochberg-corrected Mann-Whitney U test. Normalized serum level concentrations are presented in box plots. Points represent single samples. The Kruskal-Wallis test results are indicated in the plot captions. Results of the post-hoc tests are indicated in the plots. Numbers of complete observations are displayed under the plots. _

\newpage

```{r fig-hads-cov, fig.width = mm_inch(suppl$hads_cov$w), fig.height = mm_inch(suppl$hads_cov$h), fig.cap = 'Association of inflammatory markers with SARS-CoV-2 infection convalescence in the SIMMUN cohort. Association of neurotransmitter precursor aminoacids and their decay products with anti-SARS-CoV-2 antibody response in the SIMMUN cohort.'}

suppl$hads_cov$plot

```

__Supplementary Figure S\@ref(fig:fig-hads-cov). Association of inflammatory markers with SARS-CoV-2 infection convalescence in the SIMMUN cohort. Association of neurotransmitter precursor aminoacids and their decay products with anti-SARS-CoV-2 antibody response in the SIMMUN cohort.__

_(A) Serum levels of interleukin 6 (IL6), C-reactive protein (CRP), neopterin (NEO), and neutrophil-leukocyte ratio (NLR) in the SIMMUN study participants stratified by SARS-CoV-2 infections status. Statistical significance was determined by two-tailed T test. Serum levels are shown in box plots. Points represent single observations. Significance p values are displayed in the plot captions, numbers of complete observations are indicated in the plot X axes._

_(B) Correlation of TRP, KYN, KYN/TRP, PHE, TYR and PHE/TRP with the titre of immunoglobulin gamma to receptor binding domain of the SARS-CoV-2 S1 protein (anti-RBD IgG, arbitrary units [AU]) in the SIMMUN cohort (n = `r nrow(stigma$data)`) was investigated by Spearman test. Points represent single observations, blue lines with gray ribbons depict fitted generalized additive model (GAM) trends with 95% confidence intervals. Values of correlation coefficients with 95% confidence intervals and significance are indicated in the plot captions._

\newpage

```{r fig-age, fig.width = mm_inch(suppl$age$w), fig.height = mm_inch(suppl$age$h), fig.cap = 'Correlation of aminoacid neurotransmitter precursors and their decay products with age.'}

suppl$age$plot

```

__Supplementary Figure S\@ref(fig:fig-age). Correlation of aminoacid neurotransmitter precursors and their decay products with age.__

_Correlation of tryptophan (TRP), kynurenine (KYN), kynurenine/tryptophan ratio (KYN/TRP), phenylalanine (PHE), tyrosine (TYR) and phenylalanine/tyrosine ratio (PHE/TRP) with age in the SIMMUN cohort was investigated by Pearson test. Points represent single observations, blue lines with gray ribbons depict fitted linear trends with 95% confidence intervals. Values of correlation coefficients with 95% confidence intervals, significance and numbers of complete observations are indicated in the plot captions._

\newpage

```{r fig-gender, fig.width = mm_inch(suppl$gender$w), fig.height = mm_inch(suppl$gender$h), fig.cap = 'Levels of aminoacid neurotransmitter precursors and their decay products in females and males.'}

suppl$gender$plot

```

__Supplementary Figure S\@ref(fig:fig-gender). Levels of aminoacid neurotransmitter precursors and their decay products in females and males.__

_Serum concentrations of tryptophan (TRP), kynurenine (KYN), kynurenine/tryptophan ratio (KYN/TRP), phenylalanine (PHE), tyrosine (TYR) and phenylalanine/tyrosine ratio (PHE/TRP) in female and male participants of the SIMMUN study. Statistical significance was assessed by two-tailed T test. Serum levels are shown in box plots. Points represent single observations. Significance p values are displayed in the plot captions. Numbers of complete observation are indicated in the plot X axes._

\newpage

```{r fig-fit-stats, fig.width = mm_inch(suppl$fit_stats$w), fig.height = mm_inch(suppl$fit_stats$h), fig.cap = 'Root mean square error and R-squared statistics for multi-parameter linear models of aminoacid neurotransmitter precursors and their decay products in the SIMMUN cohort'}

suppl$fit_stats$plot

```

__Supplementary Figure S\@ref(fig:fig-fit-stats). Root mean square error and R-squared statistics for multi-parameter linear models of aminoacid neurotransmitter precursors and their decay products in the SIMMUN cohort.__

_Multi-parameter linear regression models of serum levels of tryptophan (TRP), kynurenine (KYN), kynurenine/tryptophan ratio (KYN/TRP), phenylalanine (PHE), tyrosine (TYR) and phenylalanine/tyrosine ratio (PHE/TRP) in the SIMMUN models were optimized by backwards elimination and their reproducibility was tested by repeated cross-validation (CV, 50 repeats, 10 folds). Values of root mean square error (RMSE, A) and $R^2$ (B) in the training data set and CV are plotted._

\newpage

# References